### 2018-05-22 manuscript/appendix_A.Rmd
update appendix A script

---
title: "Towards an Ecological Trait-data Standard - Appendix A"
author:
- name: Florian D. Schneider \footnotemark[1] \footnotemark[2] \footnotemark[3]
- name: Malte Jochum \footnotemark[4]
- name: Gaetane Le Provost \footnotemark[3]
- name: Andreas Ostrowski \footnotemark[5]
- name: Caterina Penone \footnotemark[4]
- name: David Fichtmüller \footnotemark[5]
- name: Anton Güntsch \footnotemark[5]
- name: Pete Manning \footnotemark[3]
- name: Birgitta König-Ries\footnotemark[6]
- name: Nadja K. Simons \footnotemark[7]
fontsize: 12pt
geometry: margin=0.5in
classoption: landscape
output:
  pdf_document:
    fig_caption: yes
    fig_crop: no
    keep_tex: yes
    template: pretty.template
  html_document:
    fig_caption: yes
    fig_width: 4
    highlight: null
    keep_md: yes
    theme: united
  word_document:
    fig_caption: yes
    fig_width: 6
---

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\footnotetext[1]{Corresponding author: florian.dirk.schneider@gmail.com}
\footnotetext[2]{Institute of Linguistics and Literary Studies, Technische Universität Darmstadt, Darmstadt, Germany}
\footnotetext[3]{Senckenberg Biodiversity and Climate Research Centre (BiK-F) Frankfurt am Main, Germany}
\footnotetext[4]{Institute of Plant Sciences, University of Bern, Bern, Switzerland}
\footnotetext[5]{Botanic Garden and Botanical Museum Berlin, Freie Universität Berlin, Berlin, Germany}
\footnotetext[6]{Department of Mathematics and Computer Science, Friedrich-Schiller-Universität Jena, Jena, Germany}
\footnotetext[7]{Department for Ecology and Ecosystem Management, Center for Life and Food Sciences Weihenstephan, Technische Universität München, Freising, Germany}

\newpage

```{r, echo = FALSE, warning = FALSE}
options(knitr.table.format = "latex")
library(kableExtra)
gg <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTfQOIHjJv5x_YHgq14DR7VLTvycfsl2Njj2Ip4p2a5kca8AX5t5cmaial6Bg-MX9pBDOJaJfnlR7Qj/pub?gid=0&single=true&output=csv", encoding = "UTF-8")
dt <- subset(gg, select = c(datasetName, type, DOI, organismFocus, regionalFocus, NumberTaxa, NumberTraits, Responsible, CiteAs, TermsOfUse))

kable(dt, longtable=TRUE, 
      caption = paste0("List of published trait datasets and databases; static copy of the living spreadsheet at https://goo.gl/QxzfHy (access date ",
        format(Sys.time(), '%Y-%m-%d'), ")"), 
      booktabs = T) %>%
kable_styling(latex_options = c("scale_down", "repeat_header"), font_size = 6) %>%
  column_spec(1, width = "15em") %>%
  column_spec(2, width = "5em") %>%
  column_spec(3, width = "12em") %>%
  column_spec(4, width = "6em") %>%
  column_spec(5, width = "6em") %>%
  column_spec(6, width = "6em") %>%
  column_spec(8, width = "10em") %>%
  column_spec(9, width = "25em") %>%
  column_spec(10, width = "11em") 
```