### 2018-09-18 bexis_handbook/example01.Rmd

---
title: "Example 1"
output:  
  pdf_document:
    includes:
      in_header: header.tex
fontsize: 7pt
geometry: margin=1.5in
---

```

# load library:

library(traitdataform)

# load non-standardized dataset from BExIS:
# (requires access credentials and access rights)

moths <- read.service(21247, dec = ",")

# transform to observation long-table:

dataset1 <- as.traitdata(moths, 
                  taxa = "species",  # which column contains taxa?
                  traits = c("weight",   # which columns contain traits?
                             "wing_length", "wing_width", 
                             "wing_area", "wing_loading"
                             ),
                  units = c("mg", "mm", # which units do apply
                            "mm", "mm2", "mg mm-2"), 
                  keep = c(locationID = "plot") # which other cols to keep
                  )

# standardize taxonomy:
# (applies a matching of taxonomic names to GBIF backbone taxonomy)

dataset1Std <- standardize.taxonomy(dataset1)

# define a user-provided trait thesaurus:

traitlist <- as.thesaurus(
  weight = as.trait("body_weight",
       expectedUnit = "mg", valueType = "numeric",
       identifier = "http://t-sita.cesab.org/BETSI_vizInfo.jsp?trait=Body_weight"),
  wing_length = as.trait("wing_length",
       expectedUnit = "mm", valueType = "numeric",
       identifier = ""),
  wing_width = as.trait("wing_width",
       expectedUnit = "mm", valueType = "numeric",
       identifier = ""),
  wing_area = as.trait("wing_area",
       expectedUnit = "mm2", valueType = "numeric",
       identifier = "http://t-sita.cesab.org/BETSI_vizInfo.jsp?trait=Wing_surface"),
  wing_loading = as.trait("wing_loading",
       expectedUnit = "mg mm-2", valueType = "numeric",
       identifier = "")
)

# This example does not specify identifiers, i.e. URI references to definitions
# on the internet, for all traits. It would be advisable to provide such links,
# but published ontologies do not cover all possible definitions of traits to
# date. The metadata file should therefore include a definition of the traits in
# a clear and unambiguous format.

# standardize traits:

dataset1Std <- standardize.traits(dataset1Std, traitlist)

# standardize occurrence context by Exploratories Plot ID:

dataset1Std <- standardize.exploratories(dataset1Std, plots = "locationID")

# save to txt file, ready for BExIS upload: 
# (omitting row names, use tab as field delimiter, 
# point as decimal delimiter, NA as missing value)

write.table(dataset1Std, file = "mothtraitsStd.txt", 
            row.names = FALSE, 
            sep = "\t", 
            na = "NA", 
            quote = FALSE,
            dec = "."
            )

```